Supplementary plots

### 2020-11-26 Manuscript/Manuscript_intro_methods_results.Rmd
additional analysis standard trajectory

@@

\textcolor{red}{[Here you need to focus more. You need to start with the actual problem. The question you want to answer is: How to switch from exploratory to confirmatory mode. No need to go into the shortcomings, etc. In the paragraph above I would also take up the exploratory confirmatory narrative from the first paragraph and find literature that describes how to do this. Ulis paper is pretty explicit that Type I and Type II error plays a major role here. Build on this and also look whether other literature makes recommendations for this. Describe these, but make clear that there is no satisfying solution for this important problem. We will provide suggestion here how to solve this problem. I would definitively remove many of the papers describing the other shortcomings, but focus, focus, focus... We can talk about this soon. ]}
The stark contrast between the importance of preclinical research and its shortcomings demands a change to design preclinical experiments that yield robust evidence to guide decision making towards translation. Here, we consider how to do this by simulating a preclinical research trajectory. We understand such a trajectory as a series of experiments, successive in time, generating evidence to support the decision to carry an intervention forward to clinical testing. Whereas these experiments may include different types of studies, we focus on studies examining the efficacy of a given intervention. The trajectory, for the purpose of this study, comprises exploratory studies as the first stage and within-lab replications as the second stage and we consider the state and quality of evidence generated throughout the first two stages only. Specifically, we examine the effect of different decision criteria and design choices on transition rates from the exploratory to the replication stage, as well as success rates after within-lab replication. 
**maybe** remove (When moving closer towards a decision regarding translation to humans, more stages are needed (for example, between-lab replications, multicenter studies, inclusion of positive and negative controls))

@@

When planning a replication study, the sample size has to be determined *a priori*. This is usually done using an effect size estimate from prior research. This can be taken from the literature or a pilot study. In both cases, the effect size estimate used is most likely inflated due to reasons outlined above. \textcolor{red}{[readers will have forgotten about these reasons already, please also cite literature here that shows an inflated effect size]} Employing an overestimated effect size for sample size calculation given a desired power level (for example, 0.8) and significance threshold (for example, .05), will yield a sample size that is too low to reliably detect effects. This is especially true if true underlying effects are small. When statistical power is low, non-replications will often denote a false negative under the assumption that an underlying effect exists. \textcolor{red}{[Under the assumption that an effect exists this will always be a false negative. rewrite to remove ambiguity]} An insufficiently powered replication is not suited to support or refute a previously made claim, as success and failure might be accounted for simply by inference errors.

>>>>

```{r, include=FALSE}
# options(tinytex.verbose = TRUE)
```

@@

Along the trajectory, we have to make sure that at a given stage (which implies that we are in a specific operating mode), our strategy is aligned with the goal that we have at this stage. At early stages, we want to find possibly working interventions. In statistical terms this means that we aim for a low Type II error. We do not want to miss effects. That means that our tests should be sensitive. This comes at the cost of specificity. We allow for false positives at this stage. 
Dirnagl (2016) [@dirnagl2016thomas]

*In exploratory mode, we allow for more false positives. This happens naturally, as we operate with low prior odds that our hypothesis is true.

* focus in exploration should rather be on proper descriptive statistics, including measures of variance and confidence intervals.

*

\textcolor{red}{[Here you need to focus more. You need to start with the actual problem. The question you want to answer is: How to switch from exploratory to confirmatory mode. No need to go into the shortcomings, etc. In the paragraph above I would also take up the exploratory confirmatory narrative from the first paragraph and find literature that describes how to do this. Ulis paper is pretty explicit that Type I and Type II error plays a major role here. Build on this and also look whether other literature makes recommendations for this. Describe these, but make clear that there is no satisfying solution for this important problem. We will provide suggestion here how to solve this problem. I would definitively remove many of the papers describing the other shortcomings, but focus, focus, focus... We can talk about this soon. ]}

The stark contrast between the importance of preclinical research and its shortcomings demands a change to design preclinical experiments that yield robust evidence to guide decision making towards translation. Here, we consider how to do this by simulating a preclinical research trajectory. We understand such a trajectory as a series of experiments, successive in time, generating evidence to support the decision to carry an intervention forward to clinical testing. Whereas these experiments may include different types of studies, we focus on studies examining the efficacy of a given intervention. The trajectory, for the purpose of this study, comprises exploratory studies as the first stage and within-lab replications as the second stage and we consider the state and quality of evidence generated throughout the first two stages only. Specifically, we examine the effect of different decision criteria and design choices on transition rates from the exploratory to the replication stage, as well as success rates after within-lab replication. 
**maybe** remove (When moving closer towards a decision regarding translation to humans, more stages are needed (for example, between-lab replications, multicenter studies, inclusion of positive and negative controls))

@@

When planning a replication study, the sample size has to be determined *a priori*. This is usually done using an effect size estimate from prior research. This can be taken from the literature or a pilot study. In both cases, the effect size estimate used is most likely inflated due to reasons outlined above. \textcolor{red}{[readers will have forgotten about these reasons already, please also cite literature here that shows an inflated effect size]} Employing an overestimated effect size for sample size calculation given a desired power level (for example, 0.8) and significance threshold (for example, .05), will yield a sample size that is too low to reliably detect effects. This is especially true if true underlying effects are small. When statistical power is low, non-replications denote false negatives under the assumption that an underlying effect exists. An insufficiently powered replication is not suited to support or refute a previously made claim, as success and failure might be accounted for simply by inference errors.

### 2020-11-26 Manuscript/internal_revision/supplemetary_plots.Rmd
additional analysis standard trajectory

```{r, PPVplot_Carneiro, echo = FALSE, fig.width = 45, fig.height = 10, out.width = "100%", fig.align = 'center'}
facet_names <- 
  c("0.3" = "SESOI = 0.3",
    "0.5" = "SESOI = 0.5",
    "0.7" = "SESOI = 0.7",
    "1" = "SESOI = 1.0")

plot_PPV <- 
  ggplot(data = outcomes_10EU,
         aes(x = trajectory, y = PPV_pop_prev)) + 
  geom_point(size = 3) + 
  facet_wrap(~ SESOI, nrow = 1, ncol = 4,  labeller = labeller(.rows = facet_names)) + 
  labs(x = "Trajectory", y = "Positive predictive value") +
  scale_x_discrete(labels = c("SESOI within CI \nSESOI",
                              "Significance \nSESOI",
                              "SESOI within CI \nStandard",
                              "Significance \nStandard")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 35)) +
  theme(axis.title.y = element_text(size = 35)) +
  theme(axis.text.x = element_text(size = 28, colour = "black")) +
  theme(axis.text.y = element_text(size = 28, colour = "black")) +
  theme(strip.text.x = element_text(size = 25, colour = "black", face = "bold")) +
  theme(strip.text.y = element_text(size = 25, colour = "black", face = "bold")) +
  theme(strip.background = element_rect(fill = "white", color = "black")) +
  theme(legend.title = element_text(size = 15, face = "bold")) +
  theme(legend.text = element_text(size = 14))

@@

@@

```{r, PPVplot_Szucs, echo = FALSE, fig.width = 45, fig.height = 10, out.width = "100%", fig.align = 'center'}
facet_names <- 
  c("0.3" = "SESOI = 0.3",
    "0.5" = "SESOI = 0.5",
    "0.7" = "SESOI = 0.7",
    "1" = "SESOI = 1.0")

plot_PPV <- 
  ggplot(data = outcomes_10EU,
         aes(x = trajectory, y = PPV_pop_prev)) + 
  geom_point(size = 3) + 
  facet_wrap(~ SESOI, nrow = 1, ncol = 4,  labeller = labeller(.rows = facet_names)) + 
  labs(x = "Trajectory", y = "Positive predictive value") +
  scale_x_discrete(labels = c("SESOI within CI \nSESOI",
                              "Significance \nSESOI",
                              "SESOI within CI \nStandard",
                              "Significance \nStandard")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 35)) +
  theme(axis.title.y = element_text(size = 35)) +
  theme(axis.text.x = element_text(size = 28, colour = "black")) +
  theme(axis.text.y = element_text(size = 28, colour = "black")) +
  theme(strip.text.x = element_text(size = 25, colour = "black", face = "bold")) +
  theme(strip.text.y = element_text(size = 25, colour = "black", face = "bold")) +
  theme(strip.background = element_rect(fill = "white", color = "black")) +
  theme(legend.title = element_text(size = 15, face = "bold")) +
  theme(legend.text = element_text(size = 14))

>>>>

```{r, PPVplot_Carneiro, echo = FALSE, out.width = "100%", fig.align = 'center'}
facet_names <- 
  c("0.3" = "SESOI = 0.3",
    "0.5" = "SESOI = 0.5",
    "0.7" = "SESOI = 0.7",
    "1" = "SESOI = 1.0")

plot_PPV <- 
  ggplot(data = outcomes_10EU,
         aes(x = trajectory, y = PPV_pop_prev)) + 
  geom_point(size = 1.5) + 
  facet_wrap(~ SESOI, nrow = 2, ncol = 2,  labeller = labeller(.rows = facet_names)) + 
  labs(x = "Trajectory", y = "Positive predictive value") +
  scale_x_discrete(labels = c("SESOI within CI \nSESOI",
                              "Significance \nSESOI",
                              "SESOI within CI \nStandard",
                              "Significance \nStandard")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 11)) +
  theme(axis.title.y = element_text(size = 11)) +
  theme(axis.text.x = element_text(size = 7, colour = "black")) +
  theme(axis.text.y = element_text(size = 10, colour = "black")) +
  theme(strip.text.x = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.text.y = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.background = element_rect(fill = "white", color = "black"))

@@

\newpage

@@

```{r, PPVplot_Szucs, echo = FALSE, out.width = "100%", fig.align = 'center'}
facet_names <- 
  c("0.3" = "SESOI = 0.3",
    "0.5" = "SESOI = 0.5",
    "0.7" = "SESOI = 0.7",
    "1" = "SESOI = 1.0")

plot_PPV <- 
  ggplot(data = outcomes_10EU,
         aes(x = trajectory, y = PPV_pop_prev)) + 
  geom_point(size = 1.5) + 
  facet_wrap(~ SESOI, nrow = 2, ncol = 2,  labeller = labeller(.rows = facet_names)) + 
  labs(x = "Trajectory", y = "Positive predictive value") +
  scale_x_discrete(labels = c("SESOI within CI \nSESOI",
                              "Significance \nSESOI",
                              "SESOI within CI \nStandard",
                              "Significance \nStandard")) +
  theme_bw() +
  theme(axis.title.x = element_text(size = 11)) +
  theme(axis.title.y = element_text(size = 11)) +
  theme(axis.text.x = element_text(size = 7, colour = "black")) +
  theme(axis.text.y = element_text(size = 10, colour = "black")) +
  theme(strip.text.x = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.text.y = element_text(size = 11, colour = "black", face = "bold")) +
  theme(strip.background = element_rect(fill = "white", color = "black"))
